---
title: "Member versus Casual"
date:  "2022-12-2"
author: "Joseph White"
output: html_notebook
---

This notebooks will try to use the `lehdr` package to get at detailed demographic data

## Install lehdr
Run this code only once.
```{r}
#devtools::install_github("jamgreen/lehdr")
```

## Packages
```{r}
library(knitr)
library(tidyverse)
library(janitor)
library(lubridate) # because we will probably see some dates
library(here) # a package I haven't taught you about before that doesn't do much, but ....
library(sf) # working with simple features - geospatial
library(tmap)
library(tidycensus)
library(lehdr)
library(rnaturalearth)
library(WDI)
library(tigris)
library(rgdal)
library(sp)
library(dsbox)
```


## Downloading Census Data
Key points
-downloading basic census data with the geography and population of DC in 2021
-turn census data wide to get population data as a column
```{r}
census_api_key("d44395e2fa101f82260fae6b845676d71f017b70")

census=get_acs(geography = "tract",
               variables = c("pop"="B01001_001"),
               state="DC", geometry = TRUE,year=2021)

census_wide = census %>% select(-moe) %>% spread(variable, estimate) %>%
  select(GEOID, pop, geometry) %>% st_transform(4326)
```

## Downloading Bike Data
Key points
-reading bike data from 2021
-selecting member status, start latitude and longitude, making latitude and longitude geospacial data
-create data sets for casual vs member riders
- create one data set where member and casual variables have a row each, find percentages of riders in each polygon
```{r}
bikeshare = read_csv("202209-capitalbikeshare-tripdata.csv") %>% clean_names()

bikeshare_start = bikeshare %>% select(member_casual, start_lng, start_lat) %>% st_as_sf(coords=c("start_lng","start_lat"), crs=st_crs(census_wide))

bikeshare_start_member = bikeshare_start %>% filter(member_casual == "member")

tot_member = nrow(bikeshare_start_member)

bikeshare_start_casual = bikeshare_start %>% filter(member_casual == "casual")

tot_casual = nrow(bikeshare_start_casual)

bike_census = census_wide %>% mutate(start_count_member = lengths(st_intersects(., bikeshare_start_member))) %>% mutate(start_count_member_perc = start_count_member/tot_member)

bike_census = bike_census %>% mutate(start_count_casual = lengths(st_intersects(., bikeshare_start_casual))) %>% mutate(start_count_casual_perc = start_count_casual/tot_casual)
```
```{r}
tot_casual
tot_member
```

# Plot Our Data
```{r}
tmap_mode("view")
tm_shape(bike_census) + tm_polygons(c("start_count_member_perc","start_count_casual_perc"))
```

## Lincoln Monument Coordinates
```{r}
linc_long <- -77.0502
linc_lat <- 38.8893
linc <- data.frame(linc_lat,linc_long) 
```

## Link Lincoln monument data with bike data
Key Points
-take bikeshare data and select status and coordinates
-create a data set for the Lincoln monument data
-join these data sets so we can do calculations on the two
```{r}
boxp1 <- bikeshare %>% select(member_casual, start_lng, start_lat) %>% sample_n(1000)

boxp <- boxp1 %>% mutate(lat_linc = start_lng) %>% mutate(long_linc = lat_linc) %>% select(lat_linc, long_linc, member_casual) %>% sample_n(1000)

boxp[,'lat_linc'] = linc_lat
boxp[,'long_linc'] = linc_long

boxp2 <- left_join(boxp, boxp1, by = 'member_casual') %>% sample_n(1000)
```

## haversine
```{r}
haversine <- function(long1, lat1, long2, lat2, round = 3) {
  # convert to radians
  long1 = long1 * pi / 180
  lat1  = lat1  * pi / 180
  long2 = long2 * pi / 180
  lat2  = lat2  * pi / 180
  
  R = 6371 # Earth mean radius in km
  
  a = sin((lat2 - lat1)/2)^2 + cos(lat1) * cos(lat2) * sin((long2 - long1)/2)^2
  d = R * 2 * asin(sqrt(a))
  
  return( round(d,round) ) # distance in km
}
```

## distance from linc
```{r}
linc_dist <- boxp2 %>% mutate(distance = haversine(start_lng, start_lat, long_linc, lat_linc, round = 4))
```

## boxplot
```{r}
df <- linc_dist %>% select(distance, member_casual) %>% filter(distance < 10)
ggplot(data = df)+
  geom_boxplot(aes(x= member_casual, y = distance)) + facet_wrap(~member_casual)
```

